Think deep for reef resilience

### 2025-07-26 paper/paper_draft.qmd
Add supplementary figure to show high daily variation in connectivity

Our findings on the importance of connectivity may differ due to the highly uncertain nature
of larval connectivity modelling [@vogt-vincent2023], the way larval flows and juvenile
recruitment are implemented in ADRIA - CoralBlox, and the magnitude of early bleaching
events decreasing the breeding capacity of strongly connected reefs. Analysing daily
connectivity data Vogt-Vincent et al. [-@vogt-vincent2023] suggest that time-mean
connectivity may not be suitable for capturing the seasonal - ecological time scale,
connectivity between locations due to the high daily variability in connectivity data. This
is supported by the connectivity data used in our investigation, which have large relative
standard deviations for connectivity between reefs over the 24 days used **figure**. Using
alternative summarisation methods that capture ecological-timescale patterns in
coral larval dispersal, such as cumulated implicit connectivity [@vogt-vincent2023],
future investigations may find increased importance of both depth and reef connectivity for
resilience, in line with [@hock2017; @torres2024]. Further improvements and research in
connectivity modelling are necessary to provide a robust comparison of all factors
influencing reef resilience.

There are multiple caveats in how coral ecosystem dynamics are modelled in this
investigation. We model coral colonies as objects with circular diameter values and track
these diameters over timesteps through growth. Due to the availability of quality future
projected data we do not consider the impacts of waves, cyclones, currents and Crown of
Thorns Starfish (CoTS) on coral colony survival and bleaching (**insert citations for waves
affecting survival and bleaching**). Uncertainty in connectivity modelling outputs,
satellite-derived bathymetry data and future thermal stress projections are likely to
influence the results obtained in this investigation. Incorporating further improvements in
connectivity modelling and understanding of the influence of climate change on reef
connectivity would improve the comparison of connectivity as an important factor for reef
resilience. The DHW data used in this investigation have not been bias corrected for the
Great Barrier Reef region and are part of a wider Australian dataset [@dixon2022] developed
using statistical downscaling. We investigate a range of future climate responses, thermal
projections and ecological parameterisations of ADRIA - CoralBlox to account for many
possible futures, and our results suggest the factors that influence reef resilience are
consistent across these scenarios.

@@

>>>>

Our findings on the importance of connectivity may differ due to the highly uncertain nature
of larval connectivity modelling [@vogt-vincent2023], the way larval flows and juvenile
recruitment are implemented in ADRIA - CoralBlox, and the magnitude of early bleaching
events decreasing the breeding capacity of strongly connected reefs. Analysing daily
connectivity data Vogt-Vincent et al. [-@vogt-vincent2023] suggest that time-mean
connectivity may not be suitable for capturing the seasonal - ecological time scale,
connectivity between locations due to the high daily variability in connectivity data. This
is supported by the connectivity data used in our investigation, which have large relative
standard deviations for connectivity between reefs over the 24 days used [@sfig-connectivity-rsd]. Using
alternative summarisation methods that capture ecological-timescale patterns in
coral larval dispersal, such as cumulated implicit connectivity [@vogt-vincent2023],
future investigations may find increased importance of both depth and reef connectivity for
resilience, in line with [@hock2017; @torres2024]. Further improvements and research in
connectivity modelling are necessary to provide a robust comparison of all factors
influencing reef resilience.

There are multiple caveats in how coral ecosystem dynamics are modelled in this
investigation. We model coral colonies as objects with circular diameter values and track
these diameters over timesteps through growth. Due to the availability of quality future
projected data we do not consider the impacts of waves, cyclones, currents and Crown of
Thorns Starfish (CoTS) on coral colony survival and bleaching [**Deaker et al**; **Massel and Done**]. Uncertainty in connectivity modelling outputs,
satellite-derived bathymetry data and future thermal stress projections are likely to
influence the results obtained in this investigation. Incorporating further improvements in
connectivity modelling and understanding of the influence of climate change on reef
connectivity would improve the comparison of connectivity as an important factor for reef
resilience. The DHW data used in this investigation have not been bias corrected for the
Great Barrier Reef region and are part of a wider Australian dataset [@dixon2022] developed
using statistical downscaling. We investigate a range of future climate responses, thermal
projections and ecological parameterisations of ADRIA - CoralBlox to account for many
possible futures, and our results suggest the factors that influence reef resilience are
consistent across these scenarios.

@@

![Heatmaps indicating reef connectivity relative standard deviation (RSD %). Reef connectivity relative
standard deviation represents the reef standard deviation connection strength as a percentage
of reef mean connection strength. Lighter colours indicate higher RSD % values.](../outputs/ADRIA_results/HighResCoralStress/figs/connectivity_rsd_percent.png){#sfig-connectivity-rsd width="`{julia} timeseries_width`"}
